---
title: "Journal (reproducible report)"
author: "Markus Knitt"
date: "2020-11-05"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```
# Data Science Fundamentals 
## First Challenge: Introduction to Tidyverse
```{r plot, fig.width=12, fig.height=7}
# Load libraries ----
library(tidyverse)
library(readxl)
library(lubridate)

# Importing Files ----
bikes_tbl <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/orderlines.xlsx")
bikeshops_tbl <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/bikeshops.xlsx") 

# Joining Data
left_join(orderlines_tbl, bikes_tbl, by = c("product.id" = "bike.id"))
bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

# Wrangling Data ----
bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
  # separate location column into city and state columns
  separate(col    = location,
           into   = c("city", "state"),
           sep    = ",") %>%
  
  # Add the total price (price * quantity) 
  mutate(total.price = price * quantity) %>%
  
  # Remove unnecessary columns
  select(order.date, city, state, total.price) %>%
  
  # 5.4 Rename columns
  set_names(names(.) %>% str_replace_all("\\.", "_"))

# Sales in each state ----
# Manipulate 

sales_by_state_tbl <- bike_orderlines_wrangled_tbl %>% 
  # Select the columns we need
    select(state, total_price ) %>%
 
  # Group the data by state 
    group_by(state) %>%
  
  # and summarize the sales
    summarize(sales = sum(total_price)) %>%
  
  # add currency format
  # mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))
# Visualize 
sales_by_state_tbl %>% 
  ggplot(aes(x = state, y = sales))+
  # Use geom_col for a bar plot
  geom_col(fill = "#2DC6D6") +
  # Adding labels to the bars
  geom_label(aes(label = sales_text)) +
  # Rotate x-axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 

  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title    = "Revenue in each State",
    subtitle = "Between 2015 and 2019",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )

# Sales by Year and State ----
# Manipulate
sales_by_year_and_state_tbl <- bike_orderlines_wrangled_tbl %>%
  
  # Select columns and extract year out of date
  select(order_date, total_price, state) %>%
  mutate(year = year(order_date)) %>%
  
  # Group by year & state and calculate revenue 
  group_by(year, state) %>%
  summarise(sales = sum(total_price)) %>%
  ungroup() %>%
  
  # add currency format
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))
# Step 2 - Visualize
sales_by_year_and_state_tbl %>%
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = state)) +

  # Rotate x-axis labels
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  
  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot
  # Add a trendline
  geom_smooth(method = "lm", se = FALSE) + # Adding a tren  
  # Facet
  facet_wrap(~ state) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title = "Revenue by Year and State",
    subtitle = "Upward trend in most states",
    fill = "State" # Changes the legend name
  )

```




## Second Challenge: Data Acquisition
### Getting some Data via the Cat Facts API
I chose the cat facts API for this, because I think cats are cool. 
```{r, eval = TRUE}
# add libraries
library(tidyverse)
library(httr)
library(jsonlite)

# finds the most popular cat facts         
resp <- GET("https://cat-fact.herokuapp.com/facts")

cat_facts <- resp %>% 
  .$content %>% 
  rawToChar() %>% 
  fromJSON()
cat_facts_tbl <-cat_facts[["all"]] %>%
  as_tibble()

cat_facts_tbl %>% head(n=10)
```
### Webscraping
For this challenge I scraped the prices and model names for all the bikes in the 'gravel' category of the bike company Rose.
Admittedly, this code is not great but I was short on time and couldn't afford to take the time to figure out how to make it more neat. My apologies... 
```{r, eval = TRUE}
# WEBSCRAPING ----
# Get the prices and model names for at least one category

# LIBRARIES ----

library(tidyverse) # Main Package - Loads dplyr, purrr, etc.
library(rvest)     # HTML Hacking & Web Scraping
library(xopen)     # Quickly opening URLs
library(jsonlite)  # converts JSON files to R objects
library(glue)      # concatenate strings
library(stringi)   # character string/text processing

# get the url for each bike in the product category "gravel"

gravel_groups_url_tbl <- read_html("https://www.rosebikes.de/fahrräder/gravel") %>%
  html_nodes(css = ".catalog-category-bikes__button") %>%
  html_attr("href") %>%
  enframe(name = "position", value = "subdirectory") %>%
  mutate(url = glue("https://www.rosebikes.de{subdirectory}")
  ) 

prices_tbl_1 <- read_html(gravel_groups_url_tbl$url[1]) %>%    
  html_nodes(css = (".catalog-category-model__price-current-value")) %>%
  html_text() %>%
  as_tibble()
model_names_tbl_1 <- read_html(gravel_groups_url_tbl$url[1]) %>%
  html_nodes(css = (".catalog-category-model__title")) %>%
  html_text() %>%
  as_tibble()

prices_tbl_2 <- read_html(gravel_groups_url_tbl$url[2]) %>%    
  html_nodes(css = (".catalog-category-model__price-current-value")) %>%
  html_text() %>%
  as_tibble()
model_names_tbl_2 <- read_html(gravel_groups_url_tbl$url[2]) %>%
  html_nodes(css = (".catalog-category-model__title")) %>%
  html_text() %>%
  as_tibble()

prices_tbl_3 <- read_html(gravel_groups_url_tbl$url[3]) %>%    
  html_nodes(css = (".catalog-category-model__price-current-value")) %>%
  html_text() %>%
  as_tibble()
model_names_tbl_3 <- read_html(gravel_groups_url_tbl$url[3]) %>%
  html_nodes(css = (".catalog-category-model__title")) %>%
  html_text() %>%
  as_tibble()

prices_tbl <- prices_tbl_1 %>%
  full_join(prices_tbl_2) %>%
  full_join(prices_tbl_3)
model_names_tbl <- model_names_tbl_1 %>%
  full_join(model_names_tbl_2) %>%
  full_join(model_names_tbl_3)

bikes_tbl <- bind_cols(model_names_tbl,prices_tbl) %>%
  rename(model = value...1, price = value...2)

head(bikes_tbl, n=10)
```
